\Huge{Machine Learning} \\ \LARGE Nature's Quest: Deep Learning Exploration of Global Plant Traits from Images and Geodata

### 2024-03-29 paper/info_sheet.tex

\paragraph{Software used} \textit{Describe briefly which software and libraries you relied on. If you built a particular algorithm from scratch, you can explain that here.}

>>>>

\paragraph{Software used} \textit{jupyter, python3.10, keras3, tf2.16, various data viz libraries, gcp compute engine to train the models, pandas for most of the data science related tasks}

### 2024-03-29 paper/paper.tex

How adaptable pre-trained models are to new tasks is also is also something that could be further explored. Kunze et al. \cite{kunze2017transfer} answered this question by looking at the weight difference distribution per layer early and late into training. What they found was that models trained on a similar task to the target domain only had a minor change in the weight distribution. This reasoning could possibly be applied to more accurately find the optimal backbone. Additionally, the use of a pre-trained model could be further explored by looking at the effect of the number of layers frozen during training.

\section{Conclusion}
To conclude we noticed that first and foremost the task of predicting plant traits is a hard one. It requires careful consideration during every step of the process from understanding the data to how a neural network architecture will learn from it. From our results, we observed that a model trained on images and ancillary geodata converges faster but does not necessarily have a higher $R^2$ than a model trained on images alone. In the final tests run these results were also observed with our image model having a higher $R^2$ than the combined model. However, both models still have a negative $R^2$ suggesting some core limitations in the models. Future work could look at the effect of different hyperparameters, model architectures, and data modifications to improve the prediction capabilities of the models.

>>>>

How adaptable pre-trained models are to new tasks is also something that could be further explored. Kunze et al. \cite{kunze2017transfer} answered this question by looking at the weight difference distribution per layer early and late into training. What they found was that models trained on a similar task to the target domain only had a minor change in the weight distribution. This reasoning could possibly be applied to more accurately find the optimal backbone. Additionally, the use of a pre-trained model could be further explored by looking at the effect of the number of layers frozen during training.

\section{Conclusion}
To conclude, we noticed that first and foremost the task of predicting plant traits is a hard one. It requires careful consideration during every step of the process from understanding the data to how a neural network architecture will learn from it. From our results, we observed that a model trained on images and ancillary geodata converges faster but does not necessarily have a higher $R^2$ than a model trained on images alone. In the final test run these results were also observed with our image model having a higher $R^2$ than the combined model. However, both models still have a negative $R^2$ suggesting some core limitations in the models. Future work could look at the effect of different hyperparameters, model architectures, and data modifications to improve the prediction capabilities of the models.
